Article Title: Warming resistant corals from the Gulf of Aqaba live close to their cold-water bleaching threshold
Article Authors: Jessica Bellworthy, Maoz Fine

Code written by Jessica Bellworthy


Load libraries, data, and view
```{r}
library(readr)
library(ggplot2)
library(car)
library(lme4)
library(predictmeans)
library(emmeans)
 
#Load data
data <- read.csv("acropora_rawdata_clean.csv")
View(data)
str(data)
colnames(data)
```


####### Data Visualization and Parametric Assumption testing #######

Variable: Chlorophyll concentration per surface area (chl_cm)
# 1. Plot variable against each fixed factor (temperature and day) to assess data for outliers and visualise homogenous variance 

```{r}
#plot
chl_cm = ggplot(data = data, aes(y = chlcm))+
  geom_boxplot()+
  geom_point(aes(x = 0, color = geno), size = 2.5)+
   facet_grid(day ~ treatment)+
  theme_bw()+
   theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

chl_cm
```


# 2. Levenes Test: test the assumption of homogenous variance

```{r}
#check homoscadecity
plot(data$chlcm)  

#Null hypothesis for Levenes: There is no significant difference in the variance between groups 
# i.e. if p = <0.05 there is significant differences in the variation and therefore violates the assumption of ANOVA
#homogeneous variances: p = >0.05
    
leveneTest(chlcm ~ day, data) 
leveneTest(chlcm ~ treatment, data)  
leveneTest(chlcm ~ geno, data)  
```


Check for normally distributed data within groups for the fixed factors day and treatment
# 3. Shapiro.test on raw data groups

```{r}
## Null hyp: the distribution of the data is not significantly different from a normal distribution i.e. p = > 0.05 the data are normal

# Day T2 - Winter phase
T2amb = subset(data, day =="T2" & data$treatment=='Amb')
View(T2amb)
hist(T2amb$chlcm)
qqnorm(T2amb$chlcm)
shapiro.test(T2amb$chlcm)

T2m3=subset(data, day=='T2' & treatment=='Min 3')
shapiro.test(T2m3$chlcm)

T2m6=subset(data, day=='T2' &treatment=='Min 6')
shapiro.test(T2m6$chlcm)


## Day T4 - Summer phase
T4amb=subset(data, day=='T4' & treatment=='Amb')
shapiro.test(T4amb$chlcm)

T4m3=subset(data, day=='T4' & treatment=='Min 3')
shapiro.test(T4m3$chlcm)

T4m6=subset(data, day=='T4' & treatment=='Min 6')
shapiro.test(T4m6$chlcm)
```


####### Parametric Statistical Analysis #######

4. Linear Mixed Effects Models. Perform for each dependent variable, with the fixed factor treatment and random factor genotype
lmerTest (gaussian distribution) and print anova(model). Check parametric assumptions on residuals

```{r}
# T2 Winter phase
T2data=subset(data, day=='T2')
View(T2data)
chlcm_lmer = lmer(chlcm ~ treatment + (1|geno), data = T2data, REML=FALSE)
summary(chlcm_lmer)
anova(chlcm_lmer)

residplot(chlcm_lmer, group = "treatment")


# T4 Summer phase
T4data=subset(data, day=='T4')
View(T4data)
chlcm_lmer_T4 = lmer(chlcm ~ treatment + (1|geno), data = T4data, REML=FALSE)
summary(chlcm_lmer_T4)
anova(chlcm_lmer_T4)

residplot(chlcm_lmer_T4, group = "treatment")
```


6. Post Hoc testing:
Multiple comparisons between treatments within each time point

```{r}
#“pairwise P-value plot” displaying all the P values in pairwise comparisons. Each comparison is associated with a vertical line segment that joins the scale positions of the two EMMs being compared, and whose horizontal position is determined by the P value of that comparison.

# T2 Winter phase
ph_chlcm_T2 <- emmeans(chlcm_lmer, ~ treatment)
summary(ph_chlcm_T2, infer = TRUE)
pwpm(ph_chlcm_T2)
plot(ph_chlcm_T2, comparisons = TRUE)


# T4 Summer phase
ph_chlcm_T4 <- emmeans(chlcm_lmer_T4, ~ treatment)
summary(ph_chlcm_T4, infer = TRUE)
pwpm(ph_chlcm_T4)
plot(ph_chlcm_T4, comparisons = TRUE)
```

This code completes the data visualization and statistical testing for the single dependent variable chlorophyll density (chlcm). Repeat the above script for each of the dependent variables to be examined.
